---
title: "HW_Sonam"
author: "sn952@scarletmail.rutgers.edu"
date: "2024-09-20"
output: html_document
---

# Import the Excel file without using the first row as column names
```{r}
library(readxl)
library(ggplot2)
library(dplyr)
empData <- read_excel("/Users/sonamrathod/Documents/Rutgers/Course/Business Forecasting/Week 2/Typical_Employee_Survey_Data.xlsx", 
                      col_names = FALSE)
```

# 1. Add Header Row
```{r}
names(empData) <- c("Age", "Gender", "Job_Satisfaction", "Job_Characteristic", "Years_With_Employer", "Promotion_Likelihood", "Decision_Participation", "Budget_Participation", "Pride_Working", "Turn_Down_Job", "Coworker_Relations")

# View the data with the new headers
head(empData)
```

# 2. Print out the 5-point summary and boxplot (numerical columns) for each column
```{r}
summary(empData$Age)
summary(empData$Years_With_Employer)
summary(empData)

# Boxplots for numerical columns (Age, Years_With_Employer)
boxplot(empData$Age, main = "Boxplot of Age", ylab = "Age", col = "lightblue")
mtext("Most employees, according to the boxplot, fall within the age range of 30 to 40 years.", side =1, line =1 )

boxplot(empData$Years_With_Employer, main = "Boxplot of Years With Employer", ylab = "Years with Employer", col = "lightgreen")
mtext("According to the boxplot, the majority of employees have a tenure of 10 years or less.", side =1, line =1)
```

# 3. Convert the columns that need to be converted to Factors
```{r}
empData$Gender <- factor(empData$Gender, levels = c(1, 2), labels = c("Male", "Female"))

empData$Job_Satisfaction <- factor(empData$Job_Satisfaction, levels = c(1, 2, 3, 4),
                                   labels = c("Very Satisfied", "Moderately Satisfied", "A Little Dissatisfied", "Very Dissatisfied"),
                                   ordered = TRUE)
empData$Job_Characteristic <- factor(empData$Job_Characteristic, 
                                     levels = c(1, 2, 3, 4, 5), 
                                     labels = c("High Income", "No Danger of Being Fired", "Flexible Hours", 
                                                "Opportunities for Advancement", "Enjoying the Work"))

empData$Promotion_Likelihood <- factor(empData$Promotion_Likelihood, levels = c(1, 2, 3, 4, 5),
                                       labels = c("Very Likely", "Likely", "Not Sure", "Unlikely", "Very Unlikely"), 
                                       ordered = TRUE)
empData$Decision_Participation <- factor(empData$Decision_Participation, 
                                         levels = c(1, 2, 3, 4), 
                                         labels = c("Always", "Much of the Time", "Sometimes", "Never"))

empData$Budget_Participation <- factor(empData$Budget_Participation, levels = c(1, 2), labels = c("Yes", "No"))
empData$Pride_Working <- factor(empData$Pride_Working, 
                                levels = c(1, 2, 3, 4),
                                labels = c("Very Proud", "Somewhat Proud", "Indifferent", "Not At All Proud"),
                                ordered = TRUE)

empData$Turn_Down_Job <- factor(empData$Turn_Down_Job, 
                                levels = c(1, 2, 3, 4, 5),
                                labels = c("Very Likely", "Likely", "Not Sure", "Unlikely", "Very Unlikely"),
                                ordered = TRUE)
empData$Coworker_Relations <- factor(empData$Coworker_Relations, 
                                     levels = c(1, 2, 3, 4, 5),
                                     labels = c("Very Good", "Good", "So So", "Bad", "Very Bad"),
                                     ordered = TRUE)
str(empData)
```

# 4. Plot the right visual for each column 
## 4.1 Pie chart for Gender 
```{r}
gender_count <- empData %>%
  count(Gender) %>%
  mutate(percentage = n / sum(n) * 100)

# Create a pie chart with percentage labels
ggplot(gender_count, aes(x = "", y = n, fill = Gender)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  labs(title = "Gender Distribution") +
  theme_void() +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  theme(legend.position = "right", plot.title = element_text(hjust = 0.5))
```

## 4.2 Bar plot for Job Satisfaction 
```{r}
ggplot(empData, aes(x = factor(Job_Satisfaction))) + 
  geom_bar(fill = "purple", color = "black") + 
  labs(title = "Job Satisfaction Levels", x = "Job Satisfaction", y = "Count", caption ="Job Satisfaction: The majority of employees are either very satisfied or moderately satisfied with their jobs") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

## 4.3 Bar plot for Important Job Characteristic
```{r}
ggplot(empData, aes(y = Job_Characteristic)) + 
  geom_bar(fill = "coral", color = "black") + 
  labs(title = "Important Job Characteristics", y = "Job Characteristic", x = "Count") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

## 4.4 Histogram for Age 
```{r}
ggplot(empData, aes(x = Age)) + 
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black") + 
  labs(title = "Age Distribution", x = "Age", y = "Count") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

## 4.5 Pie chart for Promotion Likelihood 
```{r}
budget_count <- empData %>%
  count(Promotion_Likelihood) %>%
  mutate(percentage = n / sum(n) * 100)
ggplot(budget_count, aes(x = "", y = n, fill = Promotion_Likelihood)) +
  geom_bar(stat = "identity", width = 1, color = "black", ) +
  coord_polar(theta = "y") +
  labs(title = "Promotion Expectancy") +
  theme_void() +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5), color ="brown") +
  theme(legend.position = "right")
```

More than 50% are not confident about being promotedwhich could be concerning, however, if we consider job satisfaction (4.2) & job characteristic(4.3) this is probably not an issue

## 4.6 Bar plot for Decision Participation 
```{r}
ggplot(empData, aes(x = Decision_Participation)) + 
  geom_bar(fill = "goldenrod", color = "black") + 
  labs(title = "Decision Participation Distribution", x = "Decision Participation", y = "Count") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

## 4.7 Pie chart for Budget Participation 
```{r}
budget_count <- empData %>%
  count(Budget_Participation) %>%
  mutate(percentage = n / sum(n) * 100)

# Create a pie chart with percentage labels
ggplot(budget_count, aes(x = "", y = n, fill = Budget_Participation)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  labs(title = "Budget Participation Distribution") +
  theme_void() +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5)) +
  theme(legend.position = "right")
```

## 4.8 Bar plot for Pride in Working 
```{r}
ggplot(empData, aes(x = Pride_Working)) + 
  geom_bar(fill = "seagreen", color = "black") + 
  labs(title = "Pride in Working Distribution", x = "Pride in Working", y = "Count") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

## 4.9 Bar plot for Turn Down Job for More Pay 
```{r}
ggplot(empData, aes(x = Turn_Down_Job)) + 
  geom_bar(fill = "coral", color = "black") + 
  labs(title = "Likelihood of Turning Down Job for More Pay", x = "Turn Down Job", y = "Count") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

## 4.10 Bar plot for Coworker Relations 
```{r}
ggplot(empData, aes(x = Coworker_Relations)) + 
  geom_bar(fill = "purple", color = "black") + 
  labs(title = "Coworker Relations Distribution", x = "Coworker Relations", y = "Count") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

## 4.11 Histogram for Years with Employer 
```{r}
ggplot(empData, aes(x = Years_With_Employer)) + 
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") + 
  labs(title = "Years with Employer Distribution", x = "Years with Employer", y = "Count") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

# 5. Plot scatter plots to show few relationships
## 5.1 Scatter plot for Age vs. Years with Employer
```{r}
ggplot(empData, aes(x = Age, y = Years_With_Employer)) + 
  geom_point(color = "blue", size = 2) + 
  labs(title = "Scatter Plot of Age vs. Years with Employer", x = "Age", y = "Years with Employer") +
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

This scatter plot highlights a relationship between employee age and tenure with the company. Older employees tend to have longer tenures.

>A scatter plot is primarily used to visualize the relationship between two numerical (continuous) variables. However, you can incorporate categorical data into a scatter plot by using visual properties like color, shape, or size to differentiate between categories. 

## 5.2 Relation of Age, Tenure and Gender
```{r}
ggplot(empData,aes(x = Age, y = Years_With_Employer, color = Gender))+
  geom_point(aes(shape = Gender),size = 2)+
  ggtitle("Impact of age, tenure and gender")+
  xlab("Age")+
  ylab("Years with employer")  + theme(plot.title = element_text(hjust = 0.5))
```

## 5.3 Relation of Age, Tenure and Budget Participation
```{r}
ggplot(empData,aes(x = Age, y = Years_With_Employer, color = Budget_Participation))+
  geom_point(aes(shape = Budget_Participation),size = 2)+
  ggtitle("Impact of age, tenure and budgetory decision")+
  xlab("Age")+
  ylab("Years with employer")  + theme(plot.title = element_text(hjust = 0.5))
```

## 5.4 Relation of Age, Tenure, Gender and Job Satisfaction
```{r}
ggplot(empData, aes(x = Age, y = Years_With_Employer, color =Gender)) + 
  geom_point(size = 3) + 
  facet_wrap(~ Job_Satisfaction) + 
  labs(title = "Age vs. Years with Employer Faceted by Job Satisfaction", 
       x = "Age", 
       y = "Years with Employer") + 
  theme_minimal() +
  
  # Center-align the title
  theme(plot.title = element_text(hjust = 0.5))
```

## 5.5 Relation of Age, Tenure, Gender and Job Characteristic
```{r}
ggplot(empData, aes(y = Years_With_Employer, x = Age, color = Gender))+
  geom_point(aes(shape = Gender), size = 2)+
  facet_wrap(~ Job_Characteristic) + 
  ggtitle("Relation of  Age, Tenure, Gender & Decision participation")+
  xlab("Age")+
  ylab("Years with employer") + theme(plot.title = element_text(hjust = 0.5))
```

## 5.6. Relation of Age, Tenure, Gender and Promotion Likelihood
```{r}
ggplot(empData, aes(y = Years_With_Employer, x = Age, color = Gender))+
  geom_point(aes(shape = Gender), size = 2)+
  facet_wrap(~ Promotion_Likelihood) + 
  ggtitle("Relation of  Age, Tenure, Gender & Promotion Likelihood")+
  xlab("Age")+
  ylab("Years with employer") + theme(plot.title = element_text(hjust = 0.5))
```

## 5.7 Relation of Age, Tenure, Gender and Decision participation
```{r}
ggplot(empData, aes(y = Years_With_Employer, x = Age, color = Gender))+
  geom_point(aes(shape = Gender), size = 2)+
  facet_wrap(~ Decision_Participation) + 
  ggtitle("Relation of  Age, Tenure, Gender & Decision participation")+
  xlab("Age")+
  ylab("Years with employer") + theme(plot.title = element_text(hjust = 0.5))
```

## 5.8 Relation of  Age, Tenure, Gender & Pride Working 
```{r}
ggplot(empData, aes(y = Years_With_Employer, x = Age, color = Gender))+
  geom_point(aes(shape = Gender), size = 2)+
  facet_wrap(~ Pride_Working) + 
  ggtitle("Relation of  Age, Tenure, Gender & Pride Working ")+
  xlab("Age")+
  ylab("Years with employer") + theme(plot.title = element_text(hjust = 0.5))
```

## 5.9 Relation of  Age, Tenure, Gender & Switching the job 
```{r}
ggplot(empData, aes(y = Years_With_Employer, x = Age, color = Gender))+
  geom_point(aes(shape = Gender), size = 2)+
  facet_wrap(~ Turn_Down_Job) + 
  ggtitle("Relation of  Age, Tenure, Gender & Switching the job ")+
  xlab("Age")+
  ylab("Years with employer") + theme(plot.title = element_text(hjust = 0.5))
```

## 5.10 Relation of  Age, Tenure, Gender & Coworker Relationship 
```{r}
ggplot(empData, aes(y = Years_With_Employer, x = Age, color = Gender))+
  geom_point(aes(shape = Gender), size = 2)+
  facet_wrap(~ Coworker_Relations) + 
  ggtitle("Relation of  Age, Tenure, Gender & Soworker Relationship ")+
  xlab("Age")+
  ylab("Years with employer") + theme(plot.title = element_text(hjust = 0.5))
```
